\textsf{Are large language models superhuman chemists?}

### 2024-03-21 src/tex/appendix.tex
parallel coordinates and updated plot
discuss the appendix figures

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of multiple choice questions and general questions per topic.}
    \label{fig:question_count_barplot_mcq_vs_general}
    \script{plot_statistics.py}
\end{figure}

@@

\subsection{Model performance}
We also evaluated the model performance on the entire \chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models.

@@

\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.

@@

@@

>>>>

\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.  
We see that the questions are generally complex to read.

@@

\Cref{fig:question_count_barplot_mcq_vs_general} shows that most questions in our corpus are \gls{mcq}.
A substantial fraction, in contrast to other benchmarks, is open-ended. 
\begin{figure}[htb]
    \centering
    \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
    \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of multiple choice questions and general questions per topic.}
    \label{fig:question_count_barplot_mcq_vs_general}
    \script{plot_statistics.py}
\end{figure}

@@

\subsection{Model performance}
We also evaluated the model performance on the entire \chembench corpus. 
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered completely correctly by the models. 
Note that this ranking is different from the one on the \enquote{tiny} subset.

@@

\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
The general pattern that the performance varies significantly between the different topics is also observed when the models are evaluated on the entire corpus. 
However, since some subjects are composed of questions from different sources the ranking of the models is in some instances different from the one on the \enquote{tiny} subset.

@@

\Cref{fig:parallel_coordinates_overall} shows this data in the form of a parallel coordinates plot. 
This visualization highlights the important observation that the ranking of the models highly depends on the questions they are evaluated on.
Only very broad benchmark have hence the potential to provide a comprehensive view of the capabilities of a model. However, even in those cases, the weighting of the different topics is crucial.
Hence, we believe that fine-grained analysis of model performance is crucial for the development of future benchmarks.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/parallel_coordinates_overall.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The parallel coordinates plot shows the performance of the models on the different topics of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models. }
    \label{fig:parallel_coordinates_overall}
    \script{make_parallel_coordinates_plot.py}
\end{figure}

To further investigate the performance of the models, we also compared the performance on different data sources.
Compared to topics, this is a more fine-grained analysis as topics can be composed of questions from different sources.
In \cref{fig:performance_per_topic} we see that the performance of the models varies significantly between the different data sources.
Interestingly, the performance of the models on questions sourced based on textbooks seems to better for our models than some of the semi-programmatically created tasks such as questions about the number of peaks in an \gls{nmr} spectrum.

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \label{fig:human_score_distribution}
    \caption{\textbf{Distribution of human scores.} The histogram shows the fraction of questions that have been answered completely correctly.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
\end{figure}
